library(mosaic)
library(tidyverse)

##Data Visualization: flights at ABIA

#plot of actual departure time vs. scheduled departure time
```{r echo = FALSE}

abia=ABIA

ggplot(data=ABIA)+
    geom_point(mapping=aes(x=CRSDepTime,y=DepTime,color=UniqueCarrier))

abia_from_austin <- ABIA %>%
  filter(Origin == "AUS")

abia_to_austin <- ABIA %>%
  filter(Dest == "AUS")
```


#flights departing from Austin every month
```{r echo = FALSE}
abia_dep_flights = abia_from_austin %>%
  group_by(Dest, Month)  %>%  # group the data points by month
  summarize(NotCancelled.sum = sum(Cancelled == 0))  # calculate sum of non-cancelled flights

ggplot(abia_dep_flights, aes(x=Month, y=NotCancelled.sum)) + 
  geom_density(stat='identity', fill = "#BF5700") +
  ylab("Number of Flights") +
  xlab("Month") +
  ggtitle("Flights Departing from AUS by Month") +
  facet_wrap(~ Dest)
```

#flights flying into Austin every month
```{r echo = FALSE}
abia_in_flights = abia_to_austin %>%
  group_by(Origin, Month)  %>%  # group the data points by month
  summarize(NotCancelled.sum = sum(Cancelled == 0))  # calculate sum of non-cancelled flights

ggplot(abia_in_flights, aes(x=Month, y=NotCancelled.sum)) + 
  geom_density(stat='identity', fill = "#BF5700") +
  ylab("Number of Flights") +
  xlab("Month") +
  ggtitle("Flights Departing from AUS by Month") +
  facet_wrap(~ Origin)
```


##Regression Practice

#regression/coefficient for equation
```{r echo = FALSE}

lm_creat=lm(creatclear~age,data=creatinine)
coef(lm_creat)
#rate=-0.6198*age+147.8129
#changes at rate of 0.6198
#for age 55: -0.6198(55)+147.8129 = 113.7239
```

#residual
```{r echo = FALSE}
resid_creat=resid(lm_creat)
resid_creat %>% which.min
age_data=data.frame(age=c(40,60))
predict(lm_creat,age_data)
#resid135=135-123.0203
#resid112=112-110.6240
```

#What creatinine clearance rate should we expect, on average, for 55 years old?
#Answer: 113.7239

#How does creatinine clearance rate change with age?
#Answer: rate decreases by 0.6198 mL/minutes for every year increase

#Whose creatinine clearance is higher for their age: 40/135 or 60/112
#Answer:60 year old: has a healthier rate of 1.376


##Green Building

#check if stories of building predicts rent
```{r echo = FALSE}
ggplot(data=greenbuildings)+
  geom_point(aes(x=leasing_rate,y=Rent))

lm(Rent ~ leasing_rate,data= greenbuildings)
#yes! it does; rent goes up by 13 cents every leasing_rate
```

#check green building vs. non-green building leasing_rate
```{r echo = FALSE}
ggplot(greenbuildings) + 
  geom_histogram(aes(x=leasing_rate, y=stat(density)), binwidth=10) + 
  facet_grid(green_rating~.)
```
#both buildings have higher rent with higher leasing rates

#recommended to sign buildings with higher leasing rates
  

##Milk Question

#N=(P-C)*Q
# =(P-C)*f(P); P and Q are couple

#f(P): change in quantity according to price
#Q=(alpha)(beta)^P
# y is Q (responsatory) and p is X (explanatory); make linear equation with log variables
```{r echo = FALSE}
plot(sales ~ price,data=milk)
plot(log(sales)~log(price),data=milk)
lm_ped=lm(log(sales)~log(price),data=milk)
coef(lm_ped)
```

#log(Q)=4.7-1.62*log(P)
#Q=e^4.7*P^-1.62

#N=(P-C)*(e^4.7*P^-1.62)
#N=(P-C)*(110*P^-1.62)

#Graphing
```{r echo = FALSE}
curve((x-1)*110*x^(-1.62),from = 1,to=9)
curve((x-1)*110*x^(-1.62),from = 2.5,to=2.7)
```
#optimal price
#Answer=$2.61